Waving and swaying

### 2014-08-11 Paper/grass.tex
Pages 1 and 2 half cleaned up

@@

transport of sediments, contaminants, dissolved oxygen, plant growth and biomass production etc \cite{Fonseca87}\cite{Nepf99}. 

@@

Similar phenomenon of large amplitude coherent oscillation of canopy in atmospheric flow known as ho-nami\cite{Inoue56}\cite{Raupach96}, have also been observed.

@@

Existing explanation of mo-nami invokes the existence of strong shear near the top of the canopy \cite{Ghisal02}\cite{Raupach96} (Ghisalberti 2002, Raupach 1996) due to

@@

The drag force experienced by flow due to presence of grass is believed to play a central role in the development of coherent eddies. This drag force experienced by flow depends 
primarily on flow velocity, fluid density and mean frontal area of the grass canopy. The mechanism leading to coherent eddies is believed to depend weakly on flexibility of grass
unless frequency of passage of coherent eddies is close to the natural frequency of the grass blade (Py el al 2008). Since natural frequency of a typical marine grass blades are
very small compared to the frequency of the passage of eddies observed in marine grass, we do not expect the mechanism leading to instability of flow to be affected by plant flexibility.
Therefore we replace the flexible grass with stiff structure to investigate dynamics leading to generation of coherent vortices. We use a mean field model for the coupling between the 
flow and the canopy. Drag force exerted by the grass is approximated via a 
continuous body force $\mathbf{f}=-N_g\mathbf{f_d}$ in the fluid momentum balance as

@@

$P$ the pressure, $\mu$ the viscosity and $\mathbf{g}$ the acceleration due to gravity. Drag force itself is modeled 

@@

normal to grass while $\hat{t},\hat{n}$ being unit vector along and normal to grass. We expect $C_T \ll C_N$ and take $C_T=0$ for rest of analysis. In the field such as marine
both $C_N \& N_g$ vary with distance from the bottom but we do not expect these variation to be central to the mechanism and therefore take $C_N \& N_g$ to be constants.

@@

The steady uniform flow resulting from solution of \eqref{navier-stokes} is expected to be unstable, leading to coherent eddies underlying monami. Steady state flow driven by   

@@

$-\frac{dP}{dx}+\mu\frac{d^2U}{dy^2}+\rho C_N d N_gU^2=0$ with no slip at bottom and zero shear at top of surface. A sample solution is shown in figure (1) and 
compared with mean flow observed in lab scale experiment. Flow consist of a region of approximately constant velocity

@@

profile in unvegetated region due to balance of viscous force and pressure gradient. No slip condition at bottom and continuity of shear stress at canopy top 
give rise to two boundary layer in the flow profile one near bottom and other near the tip of the grass. The boundary layer near the canopy top is identified 
as a free shear layer in the previous explanation of monami and its dependence on the grass density $N_g$ gives us a direct control of its thickness. 
In our model thickness of shear layer near top of grass is obtained by balance of viscous force 

@@

width \small $\delta \sim  H\left(\frac{\rho U_0 H}{\mu} C_N d N_g H\right)^{-1/3}$\normalsize where \small$U_0 = \frac{dP/dxH^2}{\mu}$\normalsize is scale for velocity and

@@

@@

\frac{\del u}{\del t}+U\frac{\del u}{\del x}+v\frac{\del U}{\del y} &= -\frac{1}{\rho}\frac{\del p}{\del x}+\frac{\mu}{\rho}\nabla^2u-2C_{N}dN_{g}Uu\\

@@

R\left[\del_t+\del_x U \right]\grad^2\psi - U_{yy}\psi_x = \grad^4\psi-\frac{\del}{\del_y}\left(\frac{2}{R}\frac{H^3}{\delta^3}U\psi_y\right)

@@

where $R= \frac{\rho U_0 H}{\mu}$ is Reynolds number. Seeking wave solution in form of $\left(u,v,\psi \right)= \left(\hat u, \hat v, \hat\phi \right)e^{ikx+\sigma t}$, 
we get generalized Orr-Sommerfield equation 
which is an eigenvalue problem. The eigenvalue $\sigma$ for a given $k$ provides growth rate and frequency associated with a perturbation of wavelength $2\pi/k$.

@@

&-\frac{\del}{\del y}\left(\frac{2}{R_e}\frac{H^3}{\delta^3}U\phi_y\right)

@@

\includegraphics[]{Low_drag_k}
\caption{Critical Wavenumber for different grass height as function of $f(\delta)$}
\label{Low_drag_k}

>>>>

\newcommand{\shreyas}[1]{{\bf (#1)}}

@@

transport of sediments, contaminants, dissolved oxygen, plant growth and biomass production etc \cite{Fonseca87,Nepf99}. 

@@

Similar phenomenon of large amplitude coherent oscillation of canopy in atmospheric flow known as ho-nami\cite{Inoue56,Raupach96}, have also been observed.

@@

Existing explanation of mo-nami invokes the existence of strong shear near the top of the canopy \cite{Ghisal02,Raupach96} due to

@@

The drag exerted by the grass bed on the flow is central to the hypothesized mechanism\cite{Ghisal02} of the instability underlying monami because it is directly responsible for realizing the shear layer structure of the flow. The flexibility of the grass blades, on the other hand, is considered to merely aid in visualizing the flow structure resulting from the instability. It has been demonstrate that the instability and the resulting flow structures are present even when the grass blades are replaced with rigid dowels (citation). Therefore, in the interest of deriving the simplest mathematical model for the linear instability, we consider the grass bed to form a rigid structure that exchanges momentum with the surrounding fluid but does not deform. Indeed, a more detailed calculation undertaken for terrestrial grass demonstrated (Py, et al 2006) that when the timescale of the instability is well-separated from the period of natural oscillations of the grass bed, the flow instability is similar to the instability of a shear layer.

% The drag force experienced by flow due to presence of grass is believed to play a central role in the development of coherent eddies. This drag force experienced by flow depends 
% primarily on flow velocity, fluid density and mean frontal area of the grass canopy. The mechanism leading to coherent eddies is believed to depend weakly on flexibility of grass
% unless frequency of passage of coherent eddies is close to the natural frequency of the grass blade (Py el al 2008). Since natural frequency of a typical marine grass blades are
% very small compared to the frequency of the passage of eddies observed in marine grass, we do not expect the mechanism leading to instability of flow to be affected by plant flexibility.
% Therefore we replace the flexible grass with stiff structure to investigate dynamics leading to generation of coherent vortices. 
We use a mean field model for the coupling between the flow and the canopy. Drag force exerted by the grass is approximated via a 
continuous body force $\mathbf{f}=-N_g\mathbf{f_d}$ in the fluid momentum balance \shreyas{Where did this model come from? Has such a model been used before? Citations to those needed.} as

@@

$P$ the pressure, $\mu$ the viscosity and $\mathbf{g}$ the acceleration due to gravity. The Reynolds number of the flow on the scale of a grass blade is expected to be O($10^2$-$10^3$), and hence the drag force is dominated by form drag, and is modeled 

@@

normal to grass while $\hat{t},\hat{n}$ being unit vector along and normal to grass. We expect $C_T \ll C_N$ and take $C_T=0$ for rest of analysis. In the field, both $C_N \& N_g$ vary with distance from the bottom but we do not expect these variation to be central to the mechanism and therefore take $C_N$ \& $N_g$ to be constants.

@@

The steady uniform flow $U(y)$ resulting from solution of \eqref{navier-stokes} is expected to be unstable, leading to coherent eddies underlying monami. Steady state flow driven by   

@@

$-{dP}/{dx}+\mu U''(y) +S(y) \rho C_N d N_gU^2=0$, where $S(y)=1$ in the vegetated layer $??<y<<??$ and $S(y)=0$ in the unvegetated layer $?? < y<< ??$. The flow in the laboratory experiments is better approximated by zero shear at the bottom ($y=??$) and top ($y=??$) of the fluid layer, as can be seen from the comparison between sample solutions and mean flow observed in laboratory experiment shown in Figure \ref{fig:basicflow}. Flow consist of a region of approximately constant velocity

@@

profile in unvegetated region due to balance of viscous force and pressure gradient. Continuity of flow speed and shear stress at canopy top 
give rise to a boundary layer in the flow profile near the tip of the grass. The boundary layer near the canopy top may be identified 
as a free shear layer\cite{Ghisal02} in the previous explanation of monami and its dependence on the grass density $N_g$ gives us a direct control of its thickness. 
In our model, the thickness of shear layer near top of grass is obtained by balance of viscous force 

@@

thickness $\delta \sim  H\left[({\rho U_0 H}/\mu) (C_N d N_g H)\right]^{-1/3}$\normalsize where \small$U_0 = \frac{dP/dxH^2}{\mu}$\normalsize is scale for velocity and

@@

\label{fig:basicflow}

@@

\frac{\del u}{\del t}+U\frac{\del u}{\del x}+v\frac{\del U}{\del y} &= -\frac{1}{\rho}\frac{\del p}{\del x}+\frac{\mu}{\rho}\nabla^2u-2S(y) C_{N}dN_{g}Uu\\

@@

R \left[ (\del_t+\del_x U )\grad^2\psi - U_{yy}\psi_x \right] = \grad^4\psi-\frac{\del}{\del_y}\left({2}\frac{H^3}{\delta^3}S(y) U\psi_y\right)

@@

where $R= {\rho U_0 H}/{\mu}$ is Reynolds number. Seeking wave solution in form of $\left(u,v,\psi \right)= \left(\hat u, \hat v, \hat\phi \right)e^{ikx+\sigma t}$, 
we get modified Orr-Sommerfield equation. The eigenvalue $\sigma$ for a given $k$ provides growth rate and frequency associated with a perturbation of wavelength $2\pi/k$.

@@

&-\frac{\del}{\del y}\left(\frac{2}{R_e}\frac{H^3}{\delta^3}S(y) U\phi_y\right)

@@

% \begin{figure}[htb]
% \includegraphics[]{Low_drag_k}
% \caption{Critical Wavenumber for different grass height as function of $f(\delta)$}
% \label{Low_drag_k}
% \end{figure}
